\chapter{\abstractname}

C is a widely used programming language, which is particularly popular in the implementation of operating systems and embedded system applications.
It is often used in the development of compilers, libraries and interpreters for other programming languages due to its efficiency.
C's efficiency comes from the fact that it is closer to the machine than other higher languages and it has a low overhead.
Unfortunately, some of the semantics for the language are described using the English language which makes this definition prone to ambiguities.

This work has three main goals.
First of all, to formalize the semantics of an imperative language (with pointers and arrays) which represents a deterministic subset of the C language and write an interpreter for it.
Second, to generate compilable and executable C code from the semantics.
Finally, to create a test suite and a test harness that allow us to verify that the behavior of the generated C program matches the behavior of the formalized semantics.

A small-step semantics for the language is formalized in the Isabelle/HOL theorem prover.
Next an interpreter for the language within the Isabelle/HOL environment is written, which allows for programs in the semantics to be executed.
Furthermore, this interpreter is proven to be correct with regard to the formal semantics.
In order to translate programs from the formal semantics to C code, a code generator was written.
This code generator uses the process of \textit{pretty printing} to translate the formal semantics to string representations which can then be exported to a file.
In order to verify the translation process, the final state of the generated program is compared to the expected final state to determine if they are equivalent.
Therefore, a method is defined to generate test instructions that check the equivalence condition for any program being translated to C.
A test harness is presented in the Isabelle/HOL environment which translates test instructions to C macros that are defined outside of this environment.
Finally, a test suite is presented along with example programs in the language.