\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{empty}

\vspace*{2cm}

\begin{center}
{\usekomafont{section} Acknowledgments}

\vspace{1cm}

I would like to express my gratitude to my supervisor, Prof. Tobias Nipkow, whose knowledge and passion in this area of research has been a source of inspiration and motivation since the beginning of my exchange year at the Technische Universität München when I took my first semantics course with him.
I would also like to thank him for providing the opportunity to work in this interesting field of research.

Additionally, I would like to give a special thanks to my advisor, Dr. Peter Lammich, for his constant guidance and support during this project.
He guided me throughout this project and provided me with helpful advice and knowledge.
I learned many things from him and his support made this journey a fun ride.

Furthermore, I wish to offer my heartfelt gratitude to Antti Sonkeri, Carlos Pérez and Jose Luis Jiménez for their valuable help at proofreading the multiple drafts of this work.

I would also like to express my warm regards to John Delgado and Manuel Gómez for their friendship and constant support over the last few years.

Last but not least, I want to convey my deepest gratitude and love to my family.
I owe everything to them, they are my inspiration and have always motivated me to persevere even when times get rough.
Eugenio, Gaudy, Santiago and Juana, without you this would not have been possible, thanks.

\end{center}

\vspace{1cm}


\cleardoublepage{}
