\documentclass[10pt]{article}
% \usepackage{fullpage}
\title{Recommendation Letter}

\author{Gabriela Limonta Marquez}

% \newcommand{\program}{Master's Degree Programme in Computer Science of the University of Helsinki}
\newcommand{\program}{Master's Programme in Computer, Communication and Information Sciences of Aalto University}

\begin{document}

\maketitle
\thispagestyle{empty}
\noindent
To the admissions committee for the \program.\\[3em]
%
%
Dear Madams and Sirs,\\[1em]
I am writing this reference at the request of Gabriela Limonta Marquez, who is applying for a Master's program in your university. 
I have known Gabriela for two years as her tutor and thesis supervisor.

In the concrete semantics course, Gabriela quickly got used to the theorem prover Isabelle/HOL, and the foundations of formal 
semantics of programming languages, and proved her progress by achieving a good overall mark and excellent homework results.

In her thesis on "Formalization of an Imperative Language with Procedures, Arrays and Pointers in Isabelle/HOL", 
she put the learned things to practice, and successfully implemented a prototype of the language along with a huge amount of regression tests.
Her results are currently used and extended in our research.

As her supervisor, I had the opportunity to observe Gabriela's progress and performance.
She is able to quickly understand new concepts and ideas, and then apply them efficiently 
and in a goal-directed manner to accomplish her task. She is a hard-working and responsible student, able to 
carry out a big amount of work on her own, and completing it successfully in short time.

Overall, I evaluate Gabriela as an intelligent, dependable, and dedicated person, ready to achieve her goals.
I recommend her to your master's program without reservation, and am confident she will complete successfully and with very good results.
If you have any further questions with regard to her background or qualifications, please do not hesitate to contact me.\\[1em]
%
Sincerely,

  Dr. Peter Lammich\\[.7em]
%
Technical University of Munich, Chair for Logic and Verification\\
Email address: lammich@in.tum.de\\
Phone number: +49 (89) 289-17326 

\end{document}
