\section{Introduction}

\begin{frame}
\frametitle{Motivation}
\framesubtitle{The C programming language}

\pause

The C language:

\pause

\begin{itemize}
\item{Closeness to the machine and low overhead allow efficiency.}
\pause
\item{Used for operating systems, embedded system applications, compilers, libraries and interpreters.}
\pause
\end{itemize}

Drawback?
\pause
\begin{itemize}
\item{Part of its semantics is defined in natural language, which makes it prone to ambiguities.}
\end{itemize}

\end{frame}


\begin{frame}
\frametitle{Motivation}
\framesubtitle{Goals of this work}

\pause
\begin{itemize}
\item{Formalization of an imperative language which covers a deterministic subset of the C semantics.}
\pause
\item{Writing of a correct interpreter inside Isabelle/HOL environment.}
\pause
\item{Generation of C code from programs in the formal semantics.}
\pause
\item{Creation of a test harness and a test suite to increase trust in the code generation process.}
\end{itemize}

\end{frame}


\begin{frame}
\frametitle{Chloe}
\framesubtitle{An imperative language, subset of C}

\begin{columns}[t]
\column{.45\textwidth}
\begin{block}{Features:}
\pause
\begin{itemize}
\item{Variables}
\pause
\item{Arrays}
\pause
\item{Pointer Arithmetic}
\pause
\item{While loop construct}
\pause
\item{If-then-else conditional construct}
\pause
\item{Functions}
\pause
\item{Dynamic memory}
\pause
\end{itemize}
\end{block}
\column{.45\textwidth}
\begin{block}{Limitations}
\begin{itemize}
\pause
\item{Proven sound and correct static type system}
\pause
\item{Concurrency}
\pause
\item{I/O operations}
\pause
\item{Goto statements}
\pause
\item{Labels}
\pause
\item{Break and continue statements}
\end{itemize}
\pause
\end{block}
\end{columns}

Chloe has enough expressive power to be Turing complete.

\end{frame}
